---
title: "Probability Distributions" 
author: "Jaqueline V Moreira" 
date: "2024-06-12" 
output: 
  # word_document: default
  html_document: default
  pdf_document: default
  # powerpoint_presentation
---

```{r "Loading packages", include=FALSE}
library(tidyverse)
library(ggpubr) #organizar e exibir vários gráficos juntos
library(ggalt) #ggplots com linhas mais suaves
library(dplyr)
library (magrittr)# compound assignment pipe-operator - %<>%
library (patchwork) # apresentar mais de um gráfico junto
# library(scales) #converter escala decimal para percentual
# library(queueing) # modelo de enfileiramento
```

```{r "color palettes", include=FALSE}
color_binomial = c("#ab9b96", "purple")
color_poisson = c("#DCAB6B","darkorange","#F28F3B","#C0BF5F","#C8553D",
                  "#BB7E5D","#BA6E6E","#8C705F", "#DE5466")
color_expo = c( "#FEA82F", "#FF2E00", "#FF4500", "#FF8C00")
color_urban= c("#2E2E3D", "#BFBFC0", "#C0BF5F", "#55556F", "#0C6291")
color_violet = c("#624763", "#F9B3D1", "#2F1847", "#C62E65", "#D63AF9", "#A63446", "#7E1946", "#FF0022")
color_candy = c("#26547C", "#EF476F", "#FFD166", "#06D6A0", "#FF0022")
```

# Binomial Distribution
**The binomial distribution is a discrete probability distribution that describes the number of successes in a fixed number of independent Bernoulli trials, each with the same probability of success. Mathematically, it is defined by the probability mass function (PMF) for k successes in n trials**

### Example

According to Statistics Canada life tables, the probability a randomly selected
90 year-old Canadian male survives for at least another year is approximately
0.82. If twenty 90-year-old Canadian males are randomly selected,

a) What is the probability exactly 18 survives for at least another year? 
b) What is the probability that last than 15 survives for another year?

Successes: The man survives for at least on year
Failure: The man dies within one year Since the sampling of men are random, it’s reasonable to consider them independent

Let x represent the number of men that survive for at least one year n = 20 p = 0.82

```{r "Binomial data frame", echo=FALSE, message=FALSE, warning=FALSE}
n = 20
p = 0.82
success <- 0:n
probability <- rep(NA,length(success))
for (i in success){
  probability[i+1] =  dbinom(i, size = n, prob = p)}
binomial_df = data.frame("Successes" = success, "Probability" = probability)
# filtrando dados para destaque no gráfico
# é necessário criar uma variavel que indique quais linhas do data frame devem ser destacadas. Para isso, utiliza-se a função mutate para criar a variável e a função case_when para selecionar as linhas de interesse.
binomial_df = binomial_df %>% mutate(x_igual_18 = case_when(Successes==18 ~ TRUE,TRUE ~ FALSE))
binomial_df = binomial_df %>% mutate(x_menor_15 = case_when(Successes<15 ~ TRUE,TRUE ~ FALSE))
```

#### a) Exactly 18
```{r "Binomial exemple 1", echo=FALSE, message=FALSE, warning=FALSE}
binomial_plot_1 = binomial_df %>% ggplot(mapping = (aes(x=Successes, y = probability))) +
  # Construindo gráficos
  geom_col(mapping = (aes(fill = x_igual_18))) +
  # Adiciona rótulos com filtro
  geom_text(aes(y = probability, label = ifelse(x_igual_18 == "TRUE", as.character(round(probability,5)), "")),
                vjust = 0.5,
                hjust = -0.5,
                angle = -90,
                color = "white",
                fontface = "bold") +
  # definindo cores
  scale_fill_manual(values = color_binomial)+
  # definindo Titulos
  labs (title = "x~Bin(20,0.82)",
       subtitle = "Binomial Distribution with n = 20 and p = 0.82",
       y = "P(x)",
       x= "Number of successes") +
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
    #legenda
    legend.position="none",
    legend.text = element_text(size=10),
    #axis
    axis.text.x = element_text(size=10),
    axis.text.y = element_text(size=10),
    #grid
    panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
binomial_plot_1
```

#### b) Last than 15
```{r "Binomial exemple 2", echo=FALSE, message=FALSE, warning=FALSE}
binomial_plot_2 = binomial_df %>% ggplot(mapping = (aes(x=Successes, y = Probability))) +
    # Construindo gráficos
  geom_col(mapping = (aes(fill = x_menor_15))) +
  # Adiciona texto
  annotate("text",
           x = 12,
           y = 0.05,
           label = round(filter(binomial_df, Successes < 15) %>% {sum(.$Probability)},5)) +
  # definindo cores
  scale_fill_manual(values = color_binomial)+
  # definindo Titulos
  labs (title = "x~Bin(20,0.82)",
       subtitle = "Binomial Distribution with n = 20 and p = 0.82",
       y = "P(x)",
       x= "Number of successes") +
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
    #legenda
    legend.position="none",
    legend.text = element_text(size=10),
    #axis
    axis.text.x = element_text(size=10),
    axis.text.y = element_text(size=10),
    #grid
    panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
binomial_plot_2
```

# Poisson Distribution
**A discrete distribution representing the number of events in a certain time
interval**
**A scenario with a constant occurring rate can be either described in terms of poisson or exponencial distribution**

```{r "Poisson lambda calcs", message=FALSE, warning=FALSE, include=FALSE}
x_axis = c(0:35)
lambda_1   =   rep(NA,length(x_axis))
lambda_1.5 =   rep(NA,length(x_axis))
lambda_2   =   rep(NA,length(x_axis))
lambda_3   =   rep(NA,length(x_axis))
lambda_5   =   rep(NA,length(x_axis))
lambda_10   =  rep(NA,length(x_axis))
lambda_15   =  rep(NA,length(x_axis))
for( i in x_axis ) {
  lambda_1[i+1]   =  dpois(x=i, lambda = 1)
  lambda_1.5[i+1] =  dpois(x=i, lambda = 1.5)
  lambda_2[i+1]   =  dpois(x=i, lambda = 2)
  lambda_3[i+1]   =  dpois(x=i, lambda = 3)
  lambda_5[i+1]   =  dpois(x=i, lambda = 5)
  lambda_10[i+1]  =  dpois(x=i, lambda = 10)
  lambda_15[i+1]  =  dpois(x=i, lambda = 15)
  }
poisson_data = data.frame("x" = x_axis,
                          "lam_1" = lambda_1,
                          "lam_1.5" = lambda_1.5,
                          "lam_2" = lambda_2,
                          "lam_3" = lambda_3,
                          "lam_5" = lambda_5,
                          "lam_10" = lambda_10,
                          "lam_15" = lambda_15)
poisson_data
```
```{r "Poisson lambda plots", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
poisson_plots = poisson_data %>% ggplot(mapping = (aes(x=x))) +
  # Construindo gráficos
  geom_line(mapping = (aes(y = lam_1, color = "1.0")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_1, color = "1.0")))+
  geom_line(mapping = (aes(y = lam_1.5, color = "1.5")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_1.5, color = "1.5")))+
  geom_line(mapping = (aes(y = lam_2, color = "2.0")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_2, color = "2.0")))+
  geom_line(mapping = (aes(y = lam_3, color = "3.0")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_3, color = "3.0")))+
  geom_line(mapping = (aes(y = lam_5, color = "5.0")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_5, color = "5.0")))+
  geom_line(mapping = (aes(y = lam_10, color = "10.0")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_10, color = "10.0")))+
  geom_line(mapping = (aes(y = lam_15, color = "15.0")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lam_15, color = "15.0")))+
  # # definindo cores
  scale_color_manual(values=color_poisson)+
  # definindo Titulos
  labs(title = "Poisson Distribution",
       subtitle = "with n = 35 and Lambda from 1 to 15",
       y = "P(x)",
       x= "Number of successes for time interval",
       color = "λ = ")+
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
        #legend
        legend.position="right",
        legend.text = element_text(size=10),
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))
poisson_plots
```

### Example

Exclusive Vines import Argentinian wine into Australia. They've begun
advertising on Facebook to direct traffic to their website where customers can
order wine online. The number of click-through sales from the ad is Poisson
distributed with a mean of 12 click-through sales per day.

μ = 12 click-through sales per day 
E(x) = μ = λ

Find the probability of getting:
a) exaclty 8 click-through sales in the first day 
b) at least 10 click-through sales in the first day

```{r "ex lambda 12 calc", message=FALSE, warning=FALSE, include=FALSE}
x_axis = c(0:35)
prob_lambda = rep(NA,length(x_axis))
for( i in x_axis ) {
  prob_lambda[i+1] =  dpois(x=i, lambda = 12)}
lambda_12 = data.frame("x_axis" = x_axis, "prob_lambda" = prob_lambda)
# filtrando dados para destaque no gráfico
lambda_12_df <- lambda_12 %>% mutate(x_igual_8 = case_when(x_axis==8 ~ TRUE, TRUE ~ FALSE),
                                     x_least_10 = case_when(x_axis>=10 ~ TRUE,TRUE ~ FALSE))
```

#### a) Exacly 8
```{r "ex lambda 12 exacly 8 plot", echo=FALSE, message=FALSE, warning=FALSE}
poisson_ex_plot_2 <- lambda_12_df %>%  ggplot(mapping = (aes(x = x_axis))) +
geom_col(mapping = (aes(x = x_axis,
                        y = prob_lambda,
                        fill = x_igual_8)),
                        width = 0.5)+
  # sobrepondo curva suavizada
  geom_xspline(mapping = (aes(y = prob_lambda)),
                              color= "#6E0D25",
                              spline_shape = -0.4)+
  geom_text(aes(y = prob_lambda, label = ifelse(x_igual_8 == "TRUE", scales::percent(round(prob_lambda,5)), "")),
                vjust = -0.7, hjust = + 1) +
  # xpline suaviza gráficos de linha, o sline_shape indicado é de -0.4.
  # quando sline_shape = 0 não há suavização
  geom_point(mapping = (aes(y = prob_lambda)),
                        color = "#6E0D25")+
  # definindo cores
  scale_fill_manual(values=color_poisson)+
  # definindo Titulos
  labs(title = "X~Poi(12)",
       subtitle = "Poisson Distribution with λ = 12",
       y = "P(x)",
       x= "Number of successes for time interval",
       fill = "exacly 8 successes") +
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size=11),
        #legend
        legend.position= "none",
        legend.text = element_blank(),
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))
poisson_ex_plot_2
```

#### b) At least 10
```{r "ex lambda 12 at least 10 plot", echo=FALSE, message=FALSE, warning=FALSE}
poisson_ex_plot_3 <- lambda_12_df %>%  ggplot(mapping = (aes(x = x_axis))) +
  geom_col(mapping = (aes(x = x_axis,
                          y = prob_lambda,
                          fill = x_least_10)), #segmentar por variavel True ou False
                      width = 0.5)+
  # Adiciona texto
  annotate("text",
           x = 20,
           y = 0.05,
           label = round(filter(lambda_12_df, x_least_10 == TRUE) %>% {sum(.$prob_lambda)},5)) +
           #  label = round(filter(binomial_df, Successes < 15) %>% {sum(.$Probability)},5)) +
  # sobrepondo curva suavizada
  geom_xspline(mapping = (aes(y = prob_lambda)),
                          color= "#6E0D25",
                          spline_shape = -0.4)+
  # xpline suaviza gráficos de linha, o sline_shape indicado é de -0.4.
  # quando sline_shape = 0 não há suavização
  geom_point(mapping = (aes(y = prob_lambda)),
                        color = "#6E0D25")+
  # definindo cores
  scale_fill_manual(values=color_poisson)+
  # definindo Titulos
  labs(title = "X~Poi(12)",
       subtitle = "Poisson Distribution with λ = 12",
       y = "P(x)",
       x= "Number of successes for time interval",
       fill = "exacly 8 successes") +
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size=11),
        #legend
        legend.position= "none",
        legend.text = element_blank(),
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))
poisson_ex_plot_3

```


# Relationship between binomial and poisson distribution

**The Binomial Distribution tends towards Poisson distribution as:**
**n tends to infinity**
**p tends to zero**
**Lambda stays constant and np = Lambda.**

**In other words, if the mean of a Binomial Distribution is equal to the rate of Lambda in a Poisson Distribution (np = Lambda) both distribution will convert to the same results as n increases and p decreases.**

**The Poisson Distribution can be used to provide a reasonable approximation to the Binomial Distribution if the sample is large (n>=50) and p is small (rare events with p<=0.05)**


```{r "binomial and poisson calc", include=FALSE}
p = 0.05 # 5%
n = c(900:3000)
x = 90
bin = rep(NA, length(n))
poi = rep(NA, length(n))
for (b in n) {
  bin[b-899] = dbinom(x=x,size = b, prob = p)
}
for (i in n) {
  poi[i-899] = dpois(x=x, lambda = i*p)
}
bin_poi = data.frame("x_axis" = n, "bin" = bin, "poi" = poi)
```
```{r "binomial and poisson plot", echo=FALSE}
bin_poi_plots = bin_poi %>% ggplot(mapping = (aes(x=x_axis))) +
  # Construindo gráficos
  geom_line(mapping = (aes(y = bin, color = "Binomial")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = bin, color = "Binomial")))+
  geom_line(mapping = (aes(y = poi, color = "Poisson")), linewidth = 0.7, linetype = 2)+
  geom_point(mapping = (aes(y = poi, color = "Poisson")))+
  # # definindo cores
  scale_color_manual(values=color_poisson)+
  # definindo Titulos
  labs(title = "Binomial and Poisson Distribution",
       subtitle = "n from 900 to 3000, p = 5% and Lambda = np",
       y = "P(x)",
       x= "Number of successes",
       color = "Distribuition")+
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
        #legend
        legend.position="top",
        legend.text = element_text(size=10),
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))
bin_poi_plots
```
```{r "binomial and poisson difference", echo=FALSE}
bin_x_90 = round(dbinom(x=x,size = 2500, prob = p),6)
poi_x_90 = round(dpois(x=x,lambda = 2500*p),6)
paste0("X~Bin(2500,0.05)")
paste0("P(X = 90) = ",bin_x_90 )
paste0(" and ")
paste0("X~Poi(2500*0.05)")
paste0("P(X = 90) = ",poi_x_90 )
paste0("Uma diferença de apenas ",round(poi_x_90 - bin_x_90,6))
```

# Binomial and Poisson example
Albinism is a rare genetic disorder that affects one in 20.000 europeans. People
with albinism produce little or none of the pigment melanin. In a random sample
of 1.000 europeans, what is the probability that exactly 2 have albnism?
In a random sampling the events can be considered independent. In this case each
trial can be considered a bernoulli trial with
p = 1/20000
In a Binomial distribution: n = 1000 x = 2
in a Poisson distribution: Lambda = n*p lambda = 1000*1/20000

```{r "binomial and poisson example", echo=FALSE}
x = 2
p = 1/20000
n = 1000
lambda = n*p
bin_x_2 = round(dbinom(x=x,size = n, prob = p),6)
poi_x_2 = round(dpois(x=x,lambda = lambda),6)
paste0("In a Binomial distribution P(X = 2) = ", bin_x_2)
paste0("In a Poisson distribution P(X = 2) = ", poi_x_2)
paste0("a difference of ", round(poi_x_2 -bin_x_2,6))
```

# Exponential Distribution

**The exponential distribution is one of the widely used continuous distributions. It is used to model the time elapsed between events.**

Poisson  = Discrete
Number of events in a time interval

Exponencial = Continuous
The time between two events

*The same event can be modeled as a Poisson or an Exponential distribution*

```{r "Expo lambda calcs", message=FALSE, warning=FALSE, include=FALSE}
lambda = c(1.5,3,6,12,18,24,30)
rate_expo = c(0.025,0.5,0.1,0.2,0.3,0.4,0.5)
x_axis = 0:35
lambda_1_5 =   rep(NA,length(x_axis))
lambda_3   =   rep(NA,length(x_axis))
lambda_6   =   rep(NA,length(x_axis))
lambda_12  =   rep(NA,length(x_axis))
lambda_18  =   rep(NA,length(x_axis))
lambda_24  =   rep(NA,length(x_axis))
lambda_30  =   rep(NA,length(x_axis))
for (i in x_axis) {
  lambda_1_5[i+1] =  dexp(x=i, rate = 0.025) # lambda de 1,5 por hora = 1.5/60min  = 0.025 por min
  lambda_3[i+1]   =  dexp(x=i, rate = 0.05)  # lambda de 3 por hora = 3/60min  = 0.05 por min
  lambda_6[i+1]   =  dexp(x=i, rate = 0.1)   # lambda de 6 por hora = 6/60min  = 0.10 por min
  lambda_12[i+1]  =  dexp(x=i, rate = 0.2)   # lambda de 12 por hora = 12/60min  = 0.2 por min
  lambda_18[i+1]  =  dexp(x=i, rate = 0.3)   # lambda de 18 por hora = 18/60min  = 0.3 por min
  lambda_24[i+1]  =  dexp(x=i, rate = 0.4)   # lambda de 24 por hora = 24/60min  = 0.4 por min
  lambda_30[i+1]  =  dexp(x=i, rate = 0.5)   # lambda de 30 por hora = 30/60min  = 0.5 por min
}
# o rate da distribuição exponencial deve ser descrito em termos da menor unidade de tempo
# por exemplo, a chegada de clientes condiz com uma distribuição poisso com lambda = 3 por hora
# dessa forma, a distribuição exponencial desses clientes será 3/60min ou 0.05

expo_data = data.frame("x" = x_axis,
                        "lambda_1_5" =  lambda_1_5,
                        "lambda_3"   =  lambda_3,
                        "lambda_6"   =  lambda_6,
                        "lambda_12"  =  lambda_12,
                        "lambda-18"  =  lambda_18,
                        "lambda_24"  =  lambda_24,
                        "lambda_30"  =  lambda_30)
expo_data
```

```{r "Expo lambda plots", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
expo_plots = expo_data %>% ggplot(mapping = (aes(x=x_axis))) +
  # Construindo gráficos
  geom_line(mapping = (aes(y = lambda_1_5, color = "1.5")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_1_5, color = "1.5")))+
  geom_line(mapping = (aes(y = lambda_3, color = "3")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_3, color = "3")))+
  geom_line(mapping = (aes(y = lambda_6, color = "6")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_6, color = "6")))+
  geom_line(mapping = (aes(y = lambda_12, color = "12")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_12, color = "12")))+
  geom_line(mapping = (aes(y = lambda_18, color = "18")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_18, color = "18")))+
  geom_line(mapping = (aes(y = lambda_24, color = "24")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_24, color = "24")))+
  geom_line(mapping = (aes(y = lambda_30, color = "30")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_30, color = "30")))+
  # # definindo cores
  scale_color_manual(values=color_violet)+
  # definindo Titulos
  labs(title = "Exponencial Distribution",
       subtitle = "with n = 35 and Lambda from 1.5 to 30",
       y = "P(x)",
       x= "Time for next arrival",
       color = "λ = ")+
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
        #legend
        legend.position="right",
        legend.text = element_text(size=10),
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))
expo_plots
```

### Example

Unique visitors arrive at a certain site by a Poisson distribution at an average rate of 3 per hour. 
The average rate of a Poisson distribution is lambda so in this case lambda is
equal to 3/hour
What is the probability that the next visitor will arrive:

a) Within 15 mins 
b) After 30 mins passes 
c) Between 40 and 60 mins

The questions reference the time interval for the next arrival NOT THE AMOUNT
OF VISITORS. In this case, the exponential distribution will be used.
Lambda of 3 means that 3 visitors arrive by hour or one visitor every 20 min on
average. To make the graph more intuitive, the scale will be built in terms of
minutes.
3 per hour = 3/60 = 0.05 clients per minute (rate in the graph)

```{r message=FALSE, warning=FALSE, "Expo lambda=3", include=FALSE}
lambda = 3
rate = 0.05
x_axis = 0:180 # minutos
expo_3 = rep(NA, length(x_axis))
for (i in x_axis){
  expo_3[i+1] = dexp(x=i, rate = rate)
}
expo_3_df = data.frame("x_axis" = x_axis, "lambda_3" = expo_3)
expo_3_df
```
```{r echo=FALSE, warning=FALSE, "Expo lambda=3 plot", paged.print=FALSE}
expo_3_plot = expo_3_df %>% ggplot(mapping = (aes(x=x_axis))) +
  # Construindo gráficos
  geom_line(mapping = (aes(y = lambda_3, color = "3")), linewidth = 0.9)+
  geom_point(mapping = (aes(y = lambda_3, color = "3")))+
  # # definindo cores
  scale_color_manual(values=color_violet)+
  # definindo Titulos
  labs(title = "Exponencial Distribution",
       subtitle = "λ = 3 exponencial rate = 0.05",
       y = "P(x)",
       x= "Time for next arrival",
       color = "λ = ")+
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
        #legend
        legend.position="none",
        legend.text = element_text(size=10),
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))
expo_3_plot
```

#### a) Within 15 min

```{r echo=FALSE, message=FALSE, warning=FALSE, "Expo lambda=3 within 10 min"}
# filtrando dados para destaque no gráfico
# é necessário criar uma variavel que indique quais linhas do data frame devem ser
# destacadas. Para isso, utiliza-se a função mutate para criar a variável e a função
# case_when para selecionar as linhas de interesse.
 expo_3_df = expo_3_df %>% mutate(x_menor_15 = case_when(x_axis<=15 ~ TRUE,TRUE ~ FALSE))
# Gerando probabilidade acumulada de acordo com o destaque anterior
 sum_ex_1 = 0
  for (i in expo_3_df$x_axis){
    if (expo_3_df$x_menor_15[i+1] == TRUE){
      sum_ex_1 = sum_ex_1+expo_3_df$lambda_3[i+1]
    }}

 expo_menor_15 = pexp(q = 15, rate = 0.05)

```

```{r echo=FALSE, warning=FALSE, "Expo lambda=3 within 15 min plot", paged.print=FALSE}

expo_3_plot1 = expo_3_df %>% ggplot(mapping = (aes(x=x_axis))) +
  # Construindo gráficos
  geom_line(mapping = (aes(y = lambda_3, color = "3")), linewidth = 0.9)+
  #geom_point(mapping = (aes(y = lambda_3, color = "3")))+
  geom_area(mapping = (aes(y = lambda_3, fill = expo_3_df$x_menor_15)))+
  annotate("text", x=15, y = 0.04, label = "57,76%")+
  # # definindo cores
  scale_color_manual(values = color_violet)+
  scale_fill_manual(values = color_violet)+
  # definindo Titulos
  labs(title = "Within 15 min",
       subtitle = "exp dist. λ = 3",
       y = "P(x)",
       x= "Time for next arrival",
       color = "λ = ")+
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
        #legend
        legend.position="none",
        legend.text = element_text(size=10),
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

expo_3_plot1

paste0("P(x<15) = ", round(expo_menor_15,4))
```

#### b) After 30 min
```{r echo=FALSE, message=FALSE, warning=FALSE, "Expo lambda=3 after 30 min"}
# filtrando dados para destaque no gráfico
# é necessário criar uma variavel que indique quais linhas do data frame devem ser
# destacadas. Para isso, utiliza-se a função mutate para criar a variável e a função
# case_when para selecionar as linhas de interesse.
 expo_3_df = expo_3_df %>% mutate(x_maior_30 = case_when(x_axis>=30 ~ TRUE,TRUE ~ FALSE))
# Gerando probabilidade acumulada de acordo com o destaque anterior

expo_maior_30 = 1 - (pexp(q = 30, rate = 0.05))

```

```{r echo=FALSE, warning=FALSE, "Expo lambda=3 after 30 min plot", paged.print=FALSE}
expo_3_plot2 = expo_3_df %>% ggplot(mapping = (aes(x=x_axis))) +
  # Construindo gráficos
  geom_line(mapping = (aes(y = lambda_3, color = "3")), linewidth = 0.9)+
  #geom_point(mapping = (aes(y = lambda_3, color = "3")))+
  geom_area(mapping = (aes(y = lambda_3, fill = x_maior_30)))+
  annotate("text", x=70, y = 0.007, label = round(expo_maior_30,4))+
  # # definindo cores
  scale_color_manual(values=color_violet)+
  scale_fill_manual(values=color_violet)+
  # definindo Titulos
  labs(title = "after 30 min",
       subtitle = "exp dist. λ = 3",
       y = "P(x)",
       x= "Time for next arrival",
       color = "λ = ")+
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
        #legend
        legend.position="none",
        legend.text = element_text(size=10),
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

expo_3_plot2

paste0("P(x>30) = ", round(expo_maior_30,4))
```


#### c) Between 40 and 60 min

```{r echo=FALSE, message=FALSE, warning=FALSE, "Expo lambda=3 between 40 and 60 min"}
# filtrando dados para destaque no gráfico
# é necessário criar uma variavel que indique quais linhas do data frame devem ser
# destacadas. Para isso, utiliza-se a função mutate para criar a variável e a função
# case_when para selecionar as linhas de interesse.
 expo_3_df = expo_3_df %>% mutate(x_menor_60 = case_when(x_axis<=60 ~ TRUE,TRUE ~ FALSE))
 expo_3_df = expo_3_df %>% mutate(x_menor_40 = case_when(x_axis<=40 ~ TRUE,TRUE ~ FALSE))
# Gerando probabilidade acumulada de acordo com o destaque anterior
expo_between_40_60 = (pexp(q = 60, rate = 0.05)) - (pexp(q = 40, rate = 0.05))

```

```{r echo=FALSE, warning=FALSE, "Expo lambda=3 between 40 and 60 min plot", paged.print=FALSE}
expo_3_60_plot = expo_3_df %>% ggplot(mapping = (aes(x=x_axis))) +
  # Construindo gráficos
  #geom_line(mapping = (aes(y = lambda_3, color = "3")), linewidth = 0.9)+
  #geom_point(mapping = (aes(y = lambda_3, color = "3")))+
  geom_area(mapping = (aes(y = lambda_3, fill = x_menor_60)))+
  annotate("text", x=60, y = 0.025, label = round(pexp(60,rate = 0.05),4))+
  # # definindo cores
  scale_color_manual(values=color_urban)+
  scale_fill_manual(values=color_urban)+
  # definindo Titulos
  labs(title = "between 40 and 60 min",
       subtitle = "exp dist. λ = 3",
       y = "P(x)",
       x= "Time for next arrival",
       color = "λ = ")+
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_text(hjust = 0.5, size=14, face="bold"),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_text(hjust = 0.5, size=11),
        #legend
        legend.position="none",
        legend.text = element_text(size=10),
        #axis
        axis.text.x = element_blank(),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

######################################

expo_3_40_plot = expo_3_df %>% ggplot(mapping = (aes(x=x_axis))) +
  # Construindo gráficos
  #geom_line(mapping = (aes(y = lambda_3, color = "3")), linewidth = 0.9)+
  #geom_point(mapping = (aes(y = lambda_3, color = "3")))+
  geom_area(mapping = (aes(y = lambda_3, fill = x_menor_40)))+
  annotate("text", x=40, y = 0.025, label = round(pexp(40,rate = 0.05),4))+
  # # definindo cores
  scale_color_manual(values=color_violet)+
  scale_fill_manual(values=color_violet)+
  # definindo Titulos
  labs(title = "Exponencial Distribution",
       subtitle = "λ = 3",
       y = "P(x)",
       x= "Time for next arrival",
       color = "λ = ")+
  # alterando aparecia
  theme_classic()+
  theme(# titles
        plot.title = element_blank(),
        axis.title.x = element_text(size=10, hjust = 0.9, face = "bold"),
        axis.title.y = element_text(size=10, hjust = 0.9, face = "bold", angle = 0),
        legend.title = element_text(size=10, face = "italic"),
        plot.subtitle = element_blank(),
        #legend
        legend.position="none",
        legend.text = element_text(size=10),
        #axis
        axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10),
        #axis.line = element_line(color = "black", size = 1, linetype = 1))
        #grid
        panel.grid.major = element_line(color = "grey", linewidth = 0.5, linetype = 2))
        #panel.grid.major.y = element_line(color = "grey", size = 0.5, linetype = 2))

######################################
expo_3_60_plot/expo_3_40_plot

paste0("P(x>40 e x<60) = 0,9502 - 0,8647")
paste0("P(x>40 e x<60) = ", round(expo_between_40_60,4))

```


